Using an SMT engine to generate Symbolic Automata

### 2019-03-04 2017-Xudong-SMT/ActaInformatica.tex
Initial , with TODO marks

\title{Using an SMT engine to generate Symbolic Automata}
\author{ Xudong Qin\autref{1}\sponsor{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai},
  Simon Bliudze\autref{2}, Eric Madelaine\autref{3}, and Min Zhang\autref{4}} 
\institute{\autlabel{1} \email{marsxd@gmail.com}\\
  \autlabel{4} \email{mzhang@sei.ecnu.edu.cn}\\
  Shanghai Key Laboratory of Trustworthy Computing, ECNU, China\par
  \autlabel{2} \email{simon.bliudze@inria.fr}\\
  INRIA Lille -- Nord Europe, 40 avenue Halley, 59650 Villeneuve d'Ascq,
France\par
  \autlabel{3} \email{eric.madelaine@inria.fr}\\
  Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia Antipolis, France
}
\date{}                                           % Activate to display a given date or no date

@@

\abstract{Open pNets are used to model the behaviour of open systems, both
synchronous or asynchronous, expressed in various calculi or
languages. They are endowed with a symbolic operational semantics in
terms of so-called ``Open Automata''. This allows us to check properties of
such systems in a compositional manner. 
We implement an algorithm computing these semantics,
building predicates expressing the synchronization conditions
between the events of the pNet sub-systems. Checking such
predicates requires symbolic reasoning over first order logics, but
also over application-specific data. We use the Z3 SMT engine to check
satisfiability of the predicates, and prune the open automaton of its
unsatisfiable transitions.
As an industrial oriented use-case, we use so-called ``architectures''
for BIP systems, that have been used in the framework of an ESA project and to specify the control software of
a nanosatellite at the EPFL Space Engineering Center. We use pNets to encode a BIP architecture
extended with explicit data, and compute its
open automaton semantics. This automaton may be used to prove
behavioural properties; we give 2 examples, a safety and a liveness
property.}
\keywords{Networks of Synchronised Automata, Symbolic Behavioral Semantics, Compositional Analysis of Software Components, SMT}

@@

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We define the open automaton generation algorithm. We
  implemented a full working prototype, within the 
  VerCors platform. In the process, we improved the
  semantics rules from~\cite{henrio:Forte2016}, and add features in
  the algorithm to deal with the full 
  model, including management of variables and assignments.
\item We implement the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the transitions
    generated by the algorithm.
\item We show the interest of this approach on an
    industry-inspired case-study, namely one architectural pattern
    extracted (and extended) from the BIP specification of a nanosatellite on-board
    software.
\end{itemize}

@@

@@

At the root of the pNet, the predicate of each OT is translated
into SMTlib assertions, and checked for satisfiability.
%
The final open automaton includes all satisfiable OTs, and the set of
reachable global states.

@@

>>>>

\title{Using an SMT engine to generate Symbolic Automata}
\author{ Xudong Qin\autref{1}\sponsor{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai},
  Simon Bliudze\autref{2}, Eric Madelaine\autref{3}, Zechen Hou\autref{1}, Yuxin Deng\autref{1}, and Min Zhang\autref{4}\todoMargin{Should we remove Min ?}} 
\institute{\autlabel{1} \email{marsxd@gmail.com}\\
  \autlabel{4} \email{mzhang@sei.ecnu.edu.cn}\\
  Shanghai Key Laboratory of Trustworthy Computing, ECNU, China\par
  \autlabel{2} \email{simon.bliudze@inria.fr}\\
  INRIA Lille -- Nord Europe, 40 avenue Halley, 59650 Villeneuve d'Ascq,
France\par
  \autlabel{3} \email{eric.madelaine@inria.fr}\\
  Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia Antipolis, France
}
\date{}                                           % Activate to display a given date or no date

@@

\abstract{Open pNets are used to model the behaviour of open systems, both
synchronous or asynchronous, expressed in various calculi or
languages. They are endowed with a symbolic operational semantics in
terms of so-called ``Open Automata''. This allows us to check properties of
such systems in a compositional manner. 
We implement an algorithm computing these semantics,
building predicates expressing the synchronization conditions
between the events of the pNet sub-systems. Checking such
predicates requires symbolic reasoning over first order logics, but
also over application-specific data. We use the Z3 SMT engine to check
satisfiability of the predicates, and prune the open automaton of its
unsatisfiable transitions.
As an industrial oriented use-case, we use so-called ``architectures''
for BIP systems, that have been used in the framework of an ESA project and to specify the control software of
a nanosatellite at the EPFL Space Engineering Center. We use pNets to encode a BIP architecture
extended with explicit data, and compute its
open automaton semantics. This automaton may be used to prove
behavioural properties; we give 2 examples, a safety and a liveness
property.
\todoMargin{Add new contrib}}
\keywords{Networks of Synchronised Automata, Symbolic Behavioral Semantics, Compositional Analysis of Software Components, SMT}

@@

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We define the open automaton generation algorithm. We
  implemented a full working prototype, within the 
  VerCors platform. In the process, we improved the
  semantics rules from~\cite{henrio:Forte2016}, and add features in
  the algorithm to deal with the full 
  model, including management of variables and assignments.
\item We implement the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the transitions
    generated by the algorithm.
    \item \ERIC{we define a strategy checking satisfiability of open transitions in the course the residual algorithm, and show that this ensures that we still generate the reachable part of the open automaton, and that what is generated is minimal with respect to the decidability of predicate decidability} 
\item We show the interest of this approach \ERIC{on various use-cases from different settings}, including an
    industry-inspired case-study, namely one architectural pattern
    extracted (and extended) from the BIP specification of a nanosatellite on-board
    software.
\end{itemize}

@@

\TODO{use a simpler running example (e.g. Lotos Enable operator ?), and move the BIP example to a dedicated section at the end}

@@

At the root of the pNet, the predicate of each OT is translated
into SMTlib assertions, and checked for satisfiability.
%
The final open automaton includes all satisfiable OTs, and the set of
reachable global states.  \TODO{Make this more precise : sat and reachability is computed only at the end...}

\subsection{Computing reachable state space on the fly}
\TODO{add a high level description of the new strategy, discuss decidability/approximation of the SAT checking, express some ``minimality'' property}

@@

\section{Use-cases}
\TODO{new section here, showing various features on different examples. Possible: Enable operator in Lotos, with (equivalent) state-based and data-based encodings; Associativity of Enable; simple communication protocol, specification and implementation; BIP FailureTimer architecture, including non-trivial ``Interval'' data-type and operators; without and with encoding of maximal progress.}
